Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}

### 2017-03-27 ident_manu.Rnw
ident threshold figure gone

\documentclass[preprint]{elsarticle}
\usepackage[paperwidth=8.5in,paperheight=11in,top=1in,bottom=1in,left=1in,right=1in]{geometry}
\usepackage{setspace}
\usepackage[colorlinks=true,allcolors=Blue]{hyperref}
\usepackage{lineno}
\usepackage{cleveref}
\usepackage{acronym}
\usepackage{paralist}
\usepackage{bm}
\usepackage{fixltx2e}
\usepackage[inline]{showlabels}

@@

An alternative view of the results in \cref{fig:identplo} can be used to demonstrate the effects of parameter selection criteria and number of parameters in the selection pool on identifiability.  \Cref{fig:percthresh} shows the percentage of identifiable parameter sets for \ac{do} using the same selection criteria in \cref{fig:identplo}, i.e., selection of parameters only within parameter categories and selection of the top sensitive parameters regardless of category.  \Cref{fig:percthresh} is similar to \cref{fig:identplo}, with the added effect of a chosen $\gamma$ threshold on identifability.  Previous studies have provided only general rules for $\gamma$ thresholds \citep{Brun01,Omlin01}, such that exact values for which parameter sets are inestimable likely vary between models and optimization methods.  As such, multiple values are shown in \cref{fig:percthresh}  given that actual thresholds could vary in practice. In general, identifiability decreased with the addition of parameters, although the rate of decrease depended on the selected threshold for $\gamma$. More conservative values for $\gamma$ (e.g., $\gamma = 5$) were more sensitive to the number of parameters in a subset, that is, identifiability decreased more quickly with the addition of parameters at lower $\gamma$ thresholds as compared to higher $\gamma$.  Notable differences in identifiability were also observed by parameter selection criteria (within categories or top parameters only), which further supports results in \cref{fig:identplo,fig:identploall}.

@@

% percent below gamma thresholds
<<percthresh, eval = T, fig.cap = 'Percent of identifiable parameter sets for \\ac{do} at different $\\gamma$ thresholds, selection criteria, and total number of parameters in the set.  Thresholds varied from $\\gamma = 2$ to $20$ such that sets with $\\gamma$ below a threshold were considered identifiable relative to the value. Plots in (a) show percent of identifiable sets by selecting parameters within categories and (b) shows percent identifiable by selecting from the top 1 through 4 parameters in all categories.  Percent identifiable was based on all sets in \\cref{fig:identplo}.', fig.height = 8, fig.width = 9, echo = F, cache = TRUE>>=
data(cat_ident_all)
data(top_ident_all)

# state variable, categories (no temp ident because only one parameter)
catnms <- c('Optics', 'Organic Matter', 'Phytoplankton', 'Zooplankton')
toplevs <- paste0('top', 1:4)
toplabs <- paste('Top', 1:4)
stts_lev <- c('O2')
stts_lab <- c('Dissolved Oxygen')

thresh <- seq(2, 20, by = 2)

# percent combos less than gamma 15, 10 by state variable, by parameter category
# temperature category not evaluated, only one paramaeter
collcatpers <- lapply(cat_ident_all[stts_lev], function(stvar){

  bystvar <- lapply(stvar[!names(stvar) %in% 'Temperature'], function(x){

    bysets <- split(x, x$N) %>% 
      lapply(function(sets){
        colls <- sets$coll
        sapply(thresh, function(x) 100 * sum(colls < x)/length(colls))

      }) %>% 
      do.call('rbind', .) %>% 
      data.frame(
        nparms = row.names(.), 
        .
      ) %>% 
      mutate(
        nparms = as.numeric(as.character(nparms))
      ) %>% 
      gather('lval', 'perc', -nparms) %>% 
      mutate(lval = factor(lval, 
        levels = paste0('X', 1:length(thresh)), 
        labels = paste0('l', thresh)
      ))
    bysets

  }) %>% 
  reshape2::melt(., id.vars = names(.[[1]])) %>% 
  rename(Category = L1)

  return(bystvar)

  }) %>% 
  reshape2::melt(., id.vars = names(.[[1]])) %>% 
  rename(`State variable` = L1) %>% 
  mutate(
    `State variable` = factor(`State variable`, levels = stts_lev, labels = stts_lab),
    Category = factor(Category, levels = catnms), 
    Threshold = gsub('^l', '', lval),
    Threshold = as.numeric(Threshold)
    )

colltoppers <- lapply(top_ident_all[stts_lev], function(stvar){

  bystvar <- lapply(stvar, function(x){

    bysets <- split(x, x$N) %>% 
      lapply(function(sets){
        colls <- sets$coll
        sapply(thresh, function(x) 100 * sum(colls < x)/length(colls))

      }) %>% 
      do.call('rbind', .) %>% 
      data.frame(
        nparms = row.names(.), 
        .
      ) %>% 
      mutate(
        nparms = as.numeric(as.character(nparms))
      ) %>% 
      gather('lval', 'perc', -nparms) %>% 
      mutate(lval = factor(lval, 
        levels = paste0('X', 1:length(thresh)), 
        labels = paste0('l', thresh)
      ))
    bysets

  }) %>% 
  reshape2::melt(., id.vars = names(.[[1]])) %>% 
  rename(Category = L1) %>% 
  filter(!Category %in% 'top5')

  return(bystvar)

  }) %>% 
  reshape2::melt(., id.vars = names(.[[1]])) %>% 
  rename(`State variable` = L1) %>% 
  mutate(
    `State variable` = factor(`State variable`, levels = stts_lev, labels = stts_lab),
    Category = factor(Category, levels = toplevs, labels = toplabs), 
    Threshold = gsub('^l', '', lval),
    Threshold = as.numeric(Threshold)
    )

toplo1 <- collcatpers
toplo2 <- colltoppers

# default theme
mytheme <- theme_minimal() + 
  theme(
    axis.line.x = element_line(), 
    axis.line.y = element_line(), 
    axis.title.x = element_blank(), 
    axis.title.y = element_blank(),
    axis.ticks.x = element_line(),
    axis.ticks.y = element_line(),
    axis.ticks.length = unit(.1, "cm"), 
    legend.position = 'none'
  )

p1 <- ggplot(toplo1, aes(x = nparms, group = lval)) + 
  geom_line(aes(y = perc), colour = lncol) + 
  geom_point(aes(y = perc, fill = Threshold), shape = 21, colour = lncol, 
    size = sz
    ) + 
  scale_fill_distiller(palette = "Spectral") +
  facet_wrap(~ Category, ncol = 1) +
  mytheme +
  ggtitle('(a) By parameter category')

p2 <- ggplot(toplo2, aes(x = nparms, group = lval)) + 
  geom_line(aes(y = perc), colour = lncol) + 
  geom_point(aes(y = perc, fill = Threshold), shape = 21, colour = lncol, 
    size = sz
    ) + 
  scale_fill_distiller(palette = "Spectral") +
  facet_wrap(~ Category, ncol = 1) +
  mytheme +
  ggtitle('(b) By top parameters, all categories')

pleg <- ggplot(toplo2, aes(x = nparms, group = lval)) + 
  geom_point(aes(y = perc, fill = Threshold)) + 
  scale_fill_distiller(palette = "Spectral")
pleg <- g_legend(pleg)

grid.arrange(
  arrangeGrob(p1, p2, ncol = 2, bottom = 'Parameters in set'),
  left = 'Percent identifiable sets',
  pleg, ncol = 2, widths = c(1, 0.2)
)
@

@@

leftlab <- expression(paste('Collinearity, ', italic(gamma)))
ggplot(toplo, aes(x = N, y = coll, fill = inc)) + 
  geom_hline(yintercept = 15, linetype = 'longdash', colour = 'darkgrey') + 
  geom_hline(yintercept = 10, linetype = 'dotted', colour = 'darkgrey') + 
  geom_boxplot(outlier.size = 0.5, size = 0.5) + 
  facet_wrap(~parm, label = label_parsed, ncol = 2) + 
  theme_minimal() + 
  scale_fill_brewer(palette = 'Spectral') + 
  scale_y_continuous(leftlab, trans = 'log', breaks = c(0.1, 1, 10, 100, 1000)) +
  scale_x_discrete('Parameters in set') + 
  theme(
    legend.title = element_blank(),
    axis.line.x = element_line(),
    axis.line.y = element_line(),
    panel.grid.minor = element_blank(), 
    panel.grid.major = element_blank(), 
    legend.position = 'top', #c(0.8, 0.1),
    axis.ticks.x = element_line(),
    axis.ticks.y = element_line(),
    axis.ticks.length = unit(.1, "cm")
  )

>>>>

\documentclass[preprint]{elsarticle}
\usepackage[paperwidth=8.5in,paperheight=11in,top=1in,bottom=1in,left=1in,right=1in]{geometry}
\usepackage{setspace}
\usepackage[colorlinks=true,allcolors=Blue]{hyperref}
\usepackage{lineno}
\usepackage{cleveref}
\usepackage{acronym}
\usepackage{paralist}
\usepackage{bm}
\usepackage{fixltx2e}
% \usepackage[inline]{showlabels}

@@

@@

@@

leftlab <- expression(paste('Collinearity, ', italic(gamma)))
ggplot(toplo, aes(x = N, y = coll, fill = inc)) + 
  geom_hline(yintercept = 15, linetype = 'longdash', colour = 'darkgrey') + 
  geom_hline(yintercept = 10, linetype = 'dotted', colour = 'darkgrey') + 
  geom_boxplot(outlier.size = 0.5, size = 0.4) + 
  facet_wrap(~parm, label = label_parsed, ncol = 2) + 
  theme_minimal() + 
  scale_fill_brewer(palette = 'Spectral') + 
  scale_y_continuous(leftlab, trans = 'log', breaks = c(0.1, 1, 10, 100, 1000)) +
  scale_x_discrete('Parameters in set') + 
  theme(
    legend.title = element_blank(),
    axis.line.x = element_line(),
    axis.line.y = element_line(),
    panel.grid.minor = element_blank(), 
    panel.grid.major = element_blank(), 
    legend.position = 'top', #c(0.8, 0.1),
    axis.ticks.x = element_line(),
    axis.ticks.y = element_line(),
    axis.ticks.length = unit(.1, "cm")
  )

### 2017-03-27 ident_manu.tex
ident threshold figure gone

\usepackage{alltt}
\usepackage[paperwidth=8.5in,paperheight=11in,top=1in,bottom=1in,left=1in,right=1in]{geometry}
\usepackage{setspace}
\usepackage[colorlinks=true,allcolors=Blue]{hyperref}
\usepackage{lineno}
\usepackage{cleveref}
\usepackage{acronym}
\usepackage{paralist}
\usepackage{bm}
\usepackage{fixltx2e}
\usepackage[inline]{showlabels}

@@

\title{Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Version: Sun Mar 26 16:14:21 2017 -0500, \href{https://github.com/fawda123/identifiability/commit/ad5c4632bfb7084298ee0ce337870bbac1db811c}{ad5c4632bfb7084298ee0ce337870bbac1db811c}}

@@

% percent below gamma thresholds
\begin{figure}[!ht]

{\centering \includegraphics[width=\maxwidth]{figs/percthresh-1}

\caption{Percent of identifiable parameter sets for \ac{do} at different $\gamma$ thresholds, selection criteria, and total number of parameters in the set.  Thresholds varied from $\gamma = 2$ to $20$ such that sets with $\gamma$ below a threshold were considered identifiable relative to the value. Plots in (a) show percent of identifiable sets by selecting parameters within categories and (b) shows percent identifiable by selecting from the top 1 through 4 parameters in all categories.  Percent identifiable was based on all sets in \cref{fig:identplo}.}\label{fig:percthresh}
\end{figure}

>>>>

\usepackage{alltt}
\usepackage[paperwidth=8.5in,paperheight=11in,top=1in,bottom=1in,left=1in,right=1in]{geometry}
\usepackage{setspace}
\usepackage[colorlinks=true,allcolors=Blue]{hyperref}
\usepackage{lineno}
\usepackage{cleveref}
\usepackage{acronym}
\usepackage{paralist}
\usepackage{bm}
\usepackage{fixltx2e}
% \usepackage[inline]{showlabels}

@@

\title{Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Version: Mon Mar 27 13:44:31 2017 -0500, \href{https://github.com/fawda123/identifiability/commit/52c5bbb2da21e0ecce154b9907a01aab993c8937}{52c5bbb2da21e0ecce154b9907a01aab993c8937}}

@@
